\plaintitle

### 2017-04-26 statistical-quorum.tex
reworded last sentence of abstract
Adding references and results.

Even if voting turnout has reached legally required quorum, it is still a question how to evaluate if
the turnout is high enough to find results legitimate.
As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.

\begin{table*}
  \centering
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Statistical quorum}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 250,055,734 & $51.5\%$ & \\
  \end{tabular}
  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union,
  and 2016 USA presidential elections, popular vote. Simplified to only two options and without invalid votes.}\label{tab:example}
\end{table*}

@@

@@

\subsection{Dynamic Quorum}

>>>>

Even if voting turnout has reached legally required quorum, or when one is not defined,
it is still a question how to evaluate if the turnout is high enough to find results legitimate.
As an example we can look at two recent votes, Brexit, a 2016 referendum for United Kingdom's withdrawal
from the European Union, and popular vote in 2016 USA presidential elections, as seen in Table~\ref{tab:example}.
Ignoring particularities of USA presidential elections process and using a popular vote as an example,
how can we reason about legitimacy of results?

\begin{table*}
  \centering
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Statistical quorum}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without invalid votes.}\label{tab:example}
\end{table*}

%traditionally, we just claim that everyone had an option

%we can use statistical quroum, defined in section X

@@

% threshold and quorum at https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf
% how does some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate them

@@

\section{Statistical Quorum}
\label{sec:statistical-quorum}
